Modeling the near-wake of a vertical-axis cross-flow turbine with 2-D and 3-D RANS

### 2016-10-02 reviews/reviewer2-response.md

>• Improvement of figures with nomenclature, labels, axis etc.

>• More explanation of why the various transport properties are investigated and their applicability

>Additionally the title could be better, maybe add a Computational Fluid Dynamics or simulation somewhere?

## Responses to PDF comments

>Effect of this [slip boundary condition] on the results?

It didn't seem to matter much. The mean velocity fields appear to be nearly
symmetrical about the $x$--$y$ plane, but nonetheless, the slip BC is at least
closer to how a free surface would behave than a no slip BC.

>Why blank? [viscous transport on in Figure 12]

These terms aren't zero, but are extremely small.

### 2016-10-02 reviews/reviewer2-response.md

## Responses to PDF comments

>Effect of this [slip boundary condition] on the results?

It didn't seem to matter much. The mean velocity fields appear to be nearly
symmetrical about the $x$--$y$ plane, but nonetheless, the slip BC is at least
closer to how a free surface would behave than a no slip BC.

>Is this mesh [Figure 4] too coarse for adequate resolution of the flow?

It could be. For our mesh dependence study we held the topology fixed and scaled
the mesh proportionally. The mesh should be fine enough to resolve the boundary
layer since $y^+ \sim 1$, but it is a fair point that maybe separated flow
structures may be dissipated in an exaggerated way. However, exploring the mesh
topology parameter space would have been prohibitive.

>Why blank? [viscous transport in Figure 12]

### 2016-10-02 paper.tex
Clarify N_x parameter sweep

Meshes were generated using OpenFOAM's blockMesh and snappyHexMesh utilities.
Mesh topology consists of a background hexahedral mesh, which is refined in all
three directions by a factor of 2 in a rectangular region containing the turbine
and near-wake (0.9 m upstream, 1.3 m downstream, $\pm 0.9$ m cross-stream, and
$\pm 0.8$ m vertically). Cells adjacent to the turbine shaft and struts are
refined by a factor of 4, while cells adjacent to the blades are refined by a
factor of 6. To capture the boundary layer, 20 layers were added next to the
blades with an expansion ratio of 1.2. Overall mesh refinement is controlled by
a single parameter---the number of cells in the streamwise direction, $N_x$.

@@

Both the $k$--$\omega$ SST and Spalart--Allmaras RANS model cases were verified
for convergence of the turbine mean power coefficient with respect to grid
spacing and time step using 2-D domains. The grid topology was fixed, but the
number of cells per domain length was scaled proportionally, maintaining the
same background mesh cell aspect ratio. Results for this parameter sweep are
shown in Figure~\ref{fig:2d-br-verification}, from which the final number of
streamwise grid points $N_x = 70$ was chosen, which resulted in a nondimensional
wall distance $y^+ = u^* y / \nu \sim 1$, where $u^*$ is the friction velocity,
defined as $u^*=\sqrt{\tau_w / \rho}$, where $\tau_w$ is the wall shear stress.
This mesh resolution gave a total cell count of approximately $5 \times 10^4$
for the 2-D cases and 16 million for the 3-D case. Sensitivity to domain length
was assessed as well, showing a 2\% increase in mean power coefficient for the
2-D SST case with the domain extended an additional two rotor diameters
downstream.

>>>>

Meshes were generated using OpenFOAM's blockMesh and snappyHexMesh utilities.
Mesh  consists of a background hexahedral mesh, which is refined in all
three directions by a factor of 2 in a rectangular region containing the turbine
and near-wake (0.9 m upstream, 1.3 m downstream, $\pm 0.9$ m cross-stream, and
$\pm 0.8$ m vertically). Cells adjacent to the turbine shaft and struts are
refined by a factor of 4, while cells adjacent to the blades are refined by a
factor of 6. To capture the boundary layer, 20 layers were added next to the
blades with an expansion ratio of 1.2. Overall mesh refinement is controlled by
a single parameter---the number of cells in the streamwise direction, $N_x$.

@@

Both the $k$--$\omega$ SST and Spalart--Allmaras RANS model cases were verified
for convergence of the turbine mean power coefficient with respect to grid
spacing and time step using 2-D domains. The grid topology was fixed, but the
number of cells per domain length $N_x$ was scaled proportionally, maintaining
the same background mesh cell aspect ratio. Results for the $N_x$ parameter
sweep are shown in Figure~\ref{fig:2d-br-verification}, from which the final
number of streamwise grid points $N_x = 70$ was chosen, which resulted in a
nondimensional wall distance $y^+ = u^* y / \nu \sim 1$, where $u^*$ is the
friction velocity, defined as $u^*=\sqrt{\tau_w / \rho}$, where $\tau_w$ is the
wall shear stress. This mesh resolution gave a total cell count of approximately
$5 \times 10^4$ for the 2-D cases and 16 million for the 3-D case. Sensitivity
to domain length was assessed as well, showing a 2\% increase in mean power
coefficient for the 2-D SST case with the domain extended an additional two
rotor diameters downstream.
